---
title: "noaastnr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{noaastnr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

You can install the development version of this package from Github with:

```r
install.packages("devtools")
devtools::install_github("UBC-MDS/noaastnr")
```
## Load the Package

You can load the package with:

```{r setup}
library(noaastnr)
```

## Usage

Typical usage will begin with downloading the list of available weather stations in the country of interest using the `get_stations_info()` function.  A dataframe is returned which can be reviewed to find a suitable station in the area of interest.  Alternatively, the NOAA provides a [graphical interface](https://gis.ncdc.noaa.gov/maps/ncei/cdo/hourly) for exploring the available weather stations.

```{r get_station_info, warning=FALSE}
get_stations_info(country = "US")
```

After selecting a weather station number, the `get_weather_data()` function can be used to download various weather parameters for the station number and year of interest.  The following usage example downloads weather data from station number "911650-22536" for the year 2020 and saves the data to a variable called 'weather_data'.  'weather_data' will be a data frame containing a time series of the following parameters for the station and year of interest:

- air temperature (degrees Celsius)
- atmospheric pressure (hectopascals)
- wind speed (m/s)
- wind direction (angular degrees)

```{r get_weather_data, warning=FALSE}
weather_data <- get_weather_data("911650-22536", 2020)
weather_data
```

The function `plot_weather_data()` can be used to visualize a time series of any of the available weather parameters either on a mean daily or mean monthly basis.  The function returns an Altair chart object which can be saved or displayed in any environment which can render Altair objects.

```{r plot_weather_data, warning=FALSE}
# weather_data <- get_weather_data("911650-22536", 2020)
plot_weather_data(weather_data, col_name="air_temp", time_basis="monthly")
```
