---
title: "noaastnr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{noaastnr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The US National Oceanic and Atmospheric Administration (NOAA) collects
and provides access to weather data from land-based weather stations
within the US and around the world ([Land-Based Station
Data](https://www.ncdc.noaa.gov/data-access/land-based-station-data)).
One method for accessing these data is through a publically accessible
FTP site. This package allows users to easily download data from a given
station for a given year, extract several key weather parameters from
the raw data files, and visualize the variation in these parameters over
time. The weather parameters that are extracted with this package are:

  - Air Temperature (degrees Celsius)
  - Atmospheric Pressure (hectopascals)
  - Wind Speed (m/s)
  - Wind Direction (angular degrees)
  
## Package Features

  - `get_stations_info`:
      - This function downloads and cleans the data of all stations
        available at <ftp://ftp.ncei.noaa.gov/pub/data/noaa/>
  - `get_weather_data`:
      - This function loads and cleans weather data for a given NOAA
        station ID and year. It returns a dataframe containing a time
        series of air temperature, atmospheric pressure, wind speed, and
        wind direction.
  - `plot_weather_data`:
      - This function visualizes the weather station observations
        including air temperature, atmospheric pressure, wind speed, and
        wind direction changing over time.
        
        

## Installation

You can install the development version of this package from [GitHub](https://github.com/) with:

```r
install.packages("devtools")
devtools::install_github("UBC-MDS/noaastnr")
```
## Load the Package

You can load the package with:

```{r setup}
library(noaastnr)
```

## Usage

Typical usage will begin with downloading the list of available weather stations in the country of interest using the `get_stations_info()` function.  A dataframe is returned which can be reviewed to find a suitable station in the area of interest.  Alternatively, the NOAA provides a [graphical interface](https://gis.ncdc.noaa.gov/maps/ncei/cdo/hourly) for exploring the available weather stations.

```{r get_station_info, warning=FALSE}
get_stations_info(country = "US")
```

After selecting a weather station number, the `get_weather_data()` function can be used to download various weather parameters for the station number and year of interest.  The following usage example downloads weather data from station number "911650-22536" for the year 2020 and saves the data to a variable called 'weather_data'.  'weather_data' will be a data frame containing a time series of the following parameters for the station and year of interest:

- air temperature (degrees Celsius)
- atmospheric pressure (hectopascals)
- wind speed (m/s)
- wind direction (angular degrees)

```{r get_weather_data, warning=FALSE}
weather_data <- get_weather_data("911650-22536", 2020)
weather_data
```

The function `plot_weather_data()` can be used to visualize a time series of any of the available weather parameters either on a mean daily or mean monthly basis.  The function returns an Altair chart object which can be saved or displayed in any environment which can render Altair objects.

```{r plot_weather_data, warning=FALSE}
# weather_data <- get_weather_data("911650-22536", 2020)
plot_weather_data(weather_data, col_name="air_temp", time_basis="monthly")
```
